---
title: "Figures_FLUXNET"
output: html_document
date: "2025-04-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 11,  # Set the plot width in inches
  fig.height = 4,  # Set the plot height in inches
  warning = FALSE,  # Hide warnings
  message = FALSE   # Hide messages
)
my_theme <- theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = "white")
  )

library(data.table)
library(lubridate)
library(dplyr)
library(Kendall)
# working directory
data_dir = ("C:/Users/yl763/Documents/GitHub/data_FLUXNET/AMF_US-Bar_FLUXNET_SUBSET_2004-2022_4-6/")
```

# FLUXNET data
- variables quick start guide <https://fluxnet.org/data/fluxnet2015-dataset/variables-quick-start-guide/>
- data variables: <https://fluxnet.org/data/aboutdata/data-variables/>
- full set data product: https://fluxnet.org/data/fluxnet2015-dataset/fullset-data-product/ 

# Working on annual data (YY)
## read in data
```{r}
setwd(data_dir); df.YY = fread("AMF_US-Bar_FLUXNET_SUBSET_YY_2004-2022_4-6.csv")
df.YY$year = as.numeric(df.YY$TIMESTAMP)
head(df.YY$NEE_VUT_REF) # gapfilled annual sums of NEE, unit: ???
head(df.YY$NEE_VUT_REF_RANDUNC) # uncertainty of annual sum, unit: ???
```

## Figure 1: annual sums of NEE
- The random uncertainty of NEE is estimated by ...(Richardson et al, 2005)
```{r}
# create upper and lower bounds
df.YY$NEE_upper <- df.YY$NEE_VUT_REF + df.YY$NEE_VUT_REF_RANDUNC
df.YY$NEE_lower <- df.YY$NEE_VUT_REF - df.YY$NEE_VUT_REF_RANDUNC

# Normalize QC values to match NEE y-axis range
max_NEE <- max(df.YY$NEE_upper, na.rm = TRUE)
min_NEE <- min(df.YY$NEE_lower, na.rm = TRUE)
range_NEE <- max_NEE - min_NEE

df.YY$QC_scaled <- min_NEE + (df.YY$NEE_VUT_REF_QC / max(df.YY$NEE_VUT_REF_QC, na.rm = TRUE)) * range_NEE
ggplot(df.YY, aes(x = year)) +
  geom_ribbon(aes(ymin = NEE_lower, ymax = NEE_upper), fill = "blue", alpha = 0.4) + # random uncertainty 
  geom_line(aes(y = NEE_VUT_REF), color = "black", size = 1) +
  geom_smooth(aes(y = NEE_VUT_REF), method = "lm", color = "red", se = FALSE, linetype = "dashed") + # Trend line
  geom_col(aes(y = QC_scaled), fill = "orange", width = 0.3, alpha = 0.5) + # QC flag bar plot
  scale_y_continuous(
    name = expression(FCO2 ~ '('*μmol ~ m^{-2} ~ s^{-1}*')'),
    sec.axis = sec_axis(
      ~ (. - min_NEE) / range_NEE * max(df.YY$NEE_VUT_REF_QC, na.rm = TRUE),
      name = "QC Flag"
    )
  ) + my_theme
  
# The Mann-Kendall trend test is used to assess whether there is a significant monotonic trend (either increasing or decreasing) in a time series. 
MannKendall(df.YY$NEE_VUT_REF) # when P < 0.05, you can reject the null hypothesis ("no trend").
```

Where to go from here:
- Interpret the figure and include it in your group presentation.
- Plot annual sums of LE and H (add these to your group presentation).


# Aggreate half hourly data (HH) to monthly scale
## read in half hourly data
```{r}
setwd(data_dir) ; df.HH = fread(paste0("AMF_US-Bar_FLUXNET_SUBSET_HH_2004-2022_4-6.csv")) 
# FLUXNET data uses "-9999" as placeholder for missing data. Replace -9999 with NA across all columns
df.HH <- df.HH %>%
  mutate(across(everything(), ~na_if(. , -9999)))
# Timestamp column of flux data is formatted as YYYYMMDDHHMM. We need to change it to a readable timestamp format.
class(df.HH$TIMESTAMP_END)
df.HH$TIMESTAMP_END <- ymd_hm(as.character(df.HH$TIMESTAMP_END))
class(df.HH$TIMESTAMP_END)
```


## Computing monthly sums 
```{r}
lambda <- 2.45e6  # Latent heat of vaporization (J/kg)
sec_per_30min <- 30*60  # Seconds in 30 minutes

# Conversion factor for converting µmol CO2 m-2 s-1 to gC per 30 min
conversion_factor_C <- (12 * 10^(-3)) * 1800 / 1000  # (mg C per µmol) * (sec per 30 min) / (mg to g)

# Create a year and month column for data summary table
df.HH <- df.HH %>%
  mutate(year = year(TIMESTAMP_END), month = month(TIMESTAMP_END))
df.HH <- df.HH %>%
  mutate(ET = (LE_F_MDS / lambda) * sec_per_30min) # convert unit from  W m-2 to mm
# Summarize the data by year and month for NEE, GPP, RECO, and additional variables
## For NEE, GPP, and Reco, converting unit from µmolCO2 mol-1 to g C m-2 month-1
monthly_data <- df.HH %>%
  group_by(year, month) %>%
  summarise(
    NEE = sum(NEE_VUT_REF * conversion_factor_C, na.rm = TRUE), 
    GPP = -sum(GPP_DT_VUT_REF * conversion_factor_C, na.rm = TRUE),
    RECO = sum(RECO_DT_VUT_REF * conversion_factor_C, na.rm = TRUE),
    P = sum(P_F, na.rm = TRUE),
    ET = sum(ET, na.rm = TRUE),
    NETRAD = sum(NETRAD, na.rm = TRUE) / (n() * 0.5),  # Convert to W/m²
    LE = sum(LE_F_MDS, na.rm = TRUE) / (n() * 0.5),  # Convert to W/m²
    H = sum(H_F_MDS, na.rm = TRUE) / (n() * 0.5),  # Convert to W/m²
    SW_IN = sum(SW_IN_F, na.rm = TRUE) / (n() * 0.5),  # Convert to W/m²
    Tair = median(TA_F, na.rm = TRUE),  # Monthly average of air temperature
    .groups = "drop"
  ) %>%
  mutate(
    month_year = as.Date(paste(year, month, 1, sep = "-")),
    year_label_pos = as.Date(paste(year, "07", "01", sep = "-")),
    y_label_pos = max(GPP, RECO, NEE, na.rm = TRUE) * 1.2
  ) 
```
Where to go from here:
- Interpret the figure and include it in your group presentation.


## Figure 2: monthly meterological variables
```{r}
# Tair -----------------------------------------------------
monthly_data <- monthly_data %>%
  mutate(y_label_pos = max(Tair, na.rm = TRUE) * 1.2)

ggplot(monthly_data, aes(x = month_year)) +
  # Add continuous lines and dots for Tair
  geom_line(aes(y = Tair, color = "Tair"), size = 1) +
  geom_point(aes(y = Tair, color = "Tair"), size = 3) +
  
  # Add horizontal dashed line at 0°C
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size = 0.8) +
  
  # Add vertical dashed lines at the end of each year
  geom_vline(data = monthly_data %>% filter(month == 12), 
             aes(xintercept = as.numeric(month_year) + 15), 
             linetype = "dashed", color = "gray", size = 0.5) +
  
  # Fix year labels position
  geom_text(data = monthly_data %>% filter(month == 6), 
            aes(x = year_label_pos, label = year), 
            y = max(monthly_data$Tair, na.rm = TRUE) * 1.2, 
            color = "black", size = 5, fontface = "bold") +
  
  # Customize the plot
  labs(x = "Month", y = expression("Air Temperature (°C)"), title = "Monthly Air Temperature Trend") +
  scale_color_manual(values = c("Tair" = "orange")) +
  scale_x_date(date_labels = "%m", date_breaks = "6 month", expand = c(0, 0)) +  # Continuous months
  scale_y_continuous(
    breaks = seq(floor(min(monthly_data$Tair, na.rm = TRUE) / 10) * 10, 
                 ceiling(max(monthly_data$Tair, na.rm = TRUE) / 10) * 10, 
                 by = 4),
    limits = c(floor(min(monthly_data$Tair, na.rm = TRUE)) - 5, 
               ceiling(max(monthly_data$Tair, na.rm = TRUE)) + 5)
  ) + my_theme

# Incoming shortwave radiation -----------------------------------------------------
monthly_data <- monthly_data %>%
  mutate(y_label_pos = max(SW_IN, na.rm = TRUE) * 1)

ggplot(monthly_data, aes(x = month_year)) +
  # Add continuous lines and dots for SW_IN
  geom_line(aes(y = SW_IN, color = "SW_IN"), size = 1) +
  geom_point(aes(y = SW_IN, color = "SW_IN"), size = 3) +
  
  # Add horizontal dashed line at 0
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size = 0.8) +
  
  # Add vertical dashed lines at the end of each year
  geom_vline(data = monthly_data %>% filter(month == 12), 
             aes(xintercept = as.numeric(month_year) + 15), 
             linetype = "dashed", color = "gray", size = 0.5) +
  
  # Fix year labels position
  geom_text(data = monthly_data %>% filter(month == 6), 
            aes(x = year_label_pos, label = year), 
            y = max(monthly_data$SW_IN, na.rm = TRUE) * 1, 
            color = "black", size = 5, fontface = "bold") +
  
  # Customize the plot
  labs(x = "Month", y = expression("Incoming Shortwave Radiation (W/m²)"), 
       title = "Monthly Incoming Shortwave Trend") +
  scale_color_manual(values = c("SW_IN" = "blue")) +
  scale_x_date(date_labels = "%m", date_breaks = "6 month", expand = c(0, 0)) +  
  scale_y_continuous(
    breaks = seq(0, 
                 ceiling(max(monthly_data$SW_IN, na.rm = TRUE) / 10) * 10, 
                 by = 50),
    limits = c(0, ceiling(max(monthly_data$SW_IN, na.rm = TRUE)) + 5)
  ) + my_theme
```
Where to go from here:
- Interpret the figure and include it in your group presentation.
- Alternatively, plot monthly anomalies of Tair and SW_IN, and add the figures to your group presentation.

## Figure 3: Monthly NEE, GPP, RECO
```{r}
monthly_data <- monthly_data %>%
  mutate(y_label_pos = max(GPP,RECO, NEE, na.rm = TRUE) * 1)

ggplot(monthly_data, aes(x = month_year)) +
  # Add bars for GPP and RECO
  geom_bar(aes(y = GPP, fill = "GPP"), stat = "identity", position = "dodge") +
  geom_bar(aes(y = RECO, fill = "RECO"), stat = "identity", position = "dodge") +
  
  # Add continuous line and dots for NEE
  geom_line(aes(y = NEE, color = "NEE"), size = 1) +
  geom_point(aes(y = NEE, color = "NEE"), size = 3) +
  
  # Add vertical dashed lines at the end of each year
  geom_vline(data = monthly_data %>% filter(month == 12), 
             aes(xintercept = as.numeric(month_year) + 15), 
             linetype = "dashed", color = "black", size = 0.5) +
  
  # Move year labels higher
  geom_text(data = monthly_data %>% filter(month == 6), 
            aes(x = year_label_pos, y = y_label_pos, label = year), 
            color = "black", size = 5, fontface = "bold") +
  
  # Define colors for bars and lines
  scale_fill_manual(values = c("GPP" = "#00BFC4", "RECO" = "#F8766D")) +
  scale_color_manual(values = c("NEE" = "purple")) +
  
  # Scale for the left y-axis (Carbon fluxes)
  scale_y_continuous(
    name = expression("g C per month"^2),
    breaks = seq(floor(min(monthly_data$NEE, monthly_data$GPP, monthly_data$RECO, na.rm = TRUE) / 50) * 50, 
                 ceiling(max(monthly_data$NEE, monthly_data$GPP, monthly_data$RECO, na.rm = TRUE) / 50) * 50, 
                 by = 20),
    )+
  # X-axis settings
  scale_x_date(date_labels = "%m", date_breaks = "6 month", expand = c(0, 0)) +  my_theme
```


# Working on DD data (experimental)
```{r}
setwd(data_dir); df.DD = fread("AMF_US-Bar_FLUXNET_SUBSET_DD_2004-2022_4-6.csv")
names(df.DD)
# library(dplyr)
# library(lubridate)
# library(SPEI)
# 
# 
# # STEP 1: Convert TIMESTAMP to date
# df.DD <- df.DD %>%
#   mutate(
#     TIMESTAMP = ymd_hms(TIMESTAMP),
#     date = as_date(TIMESTAMP)
#   )
# 
# # STEP 2: Calculate daily Tmin, Tmax, P
# daily_temp <- df.DD %>%
#   group_by(date) %>%
#   summarise(
#     Tmin = min(TA_F, na.rm = TRUE),
#     Tmax = max(TA_F, na.rm = TRUE),
#     P = sum(P_F, na.rm = TRUE)
#   ) %>%
#   filter(is.finite(Tmin), is.finite(Tmax)) %>%
#   ungroup()
# 
# # STEP 3: Compute daily PET
# lat <- 42.5  # change this to your site latitude
# daily_temp$PET <- hargreaves(Tmin = daily_temp$Tmin, Tmax = daily_temp$Tmax, lat = lat)
# 
# # STEP 4: Compute daily water balance
# daily_temp$bal <- daily_temp$P - daily_temp$PET
# 
# # STEP 5: Aggregate to monthly water balance
# monthly_balance <- daily_temp %>%
#   mutate(year = year(date), month = month(date)) %>%
#   group_by(year, month) %>%
#   summarise(water_bal = sum(bal, na.rm = TRUE)) %>%
#   ungroup() %>%
#   filter(!is.na(year), !is.na(month), !is.na(water_bal)) %>%  # NEW: remove NA rows
#   arrange(year, month)
# 
# # STEP 6: Convert to time series (after filtering)
# bal_ts <- ts(monthly_balance$water_bal,
#              start = c(min(monthly_balance$year), min(monthly_balance$month)),
#              frequency = 12)
# 
# 
# # STEP 7: Compute 12-month SPEI
# spei_12 <- spei(bal_ts, scale = 12)
# monthly_balance$SPEI_12 <- as.numeric(spei_12$fitted)
# 
# # STEP 8: Add date column for plotting
# monthly_balance$date <- as.Date(paste(monthly_balance$year, monthly_balance$month, "15", sep = "-"))
# 
# # STEP 9: Plot the SPEI
# ggplot(monthly_balance, aes(x = date, y = SPEI_12)) +
#   geom_line(color = "#2166ac", size = 1) +
#   geom_hline(yintercept = 0, linetype = "dashed", color = "gray30") +
#   geom_hline(yintercept = c(-1, -1.5, -2), linetype = "dotted", color = "red") +
#   geom_hline(yintercept = c(1, 1.5, 2), linetype = "dotted", color = "darkgreen") +
#   labs(
#     title = "12-Month SPEI from df.DD",
#     x = "Date",
#     y = "SPEI (12-month)"
#   ) +
#   theme_minimal(base_size = 14)
```

# Optional: Qualify flags
```{r, fig.height=5, fig.width=9}
df.HH <- df.HH %>% # using half-hourly data
  mutate(
    year = year(TIMESTAMP_END),
    DOY = yday(TIMESTAMP_END) + 
      (hour(TIMESTAMP_END) + minute(TIMESTAMP_END)/60 + second(TIMESTAMP_END)/3600) / 24,
    qc_label = factor(
      NEE_VUT_REF_QC,
      levels = c(0, 1, 2, 3),
      labels = c("Measured (0)", "Good (1)", "Medium (2)", "Poor (3)")
    )
  )

#calculate the percentage of each flag
qc_summary <- df.HH %>%
  group_by(year, qc_label) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(year) %>%
  mutate(
    total = sum(count),
    percent = round(100 * count / total, 1),
    label = paste0(qc_label, ": ", percent, "%")
  )
# View(qc_summary)

# plot qc_summary by year
qc_summary %>%
  filter(year >= 2004, year <= 2022) %>%
  ggplot(aes(x = factor(year), y = percent, fill = qc_label)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(
    name = "QC Flag",
    values = c(
      "Measured (0)" = "black",
      "Good (1)" = "green4",
      "Medium (2)" = "orange",
      "Poor (3)" = "red"
    )
  ) +
  labs(
    title = "Summary of NEE Quality Flags",
    x = "Year", y = "Percentage (%)"
  ) + my_theme
  

qc_labels <- qc_summary %>%
  group_by(year) %>%
  summarise(
    qc_text = paste(label, collapse = "\n"),
    .groups = "drop"
  )

data_with_labels <- df.HH %>%
  left_join(qc_labels, by = "year")

# looking into data for one year: plot NEE with QC flags
data_with_labels %>%
  filter(year == 2010) %>%
  ggplot(aes(x = DOY, y = NEE_VUT_REF, color = qc_label)) +
  geom_point(alpha = 0.6, size = 1) +
  scale_color_manual(
    name = "QC Flag",
    values = c("Measured (0)" = "black",
               "Good (1)" = "green4",
               "Medium (2)" = "orange",
               "Poor (3)" = "red")
  ) +
  geom_text(
    data = qc_labels,
    aes(x = 12, y = 20, label = qc_text),
    inherit.aes = FALSE,
    hjust = 0,
    vjust = 1.1,
    size = 3,
    color = "gray30"
  ) +
  labs(
    title = "Half hourly FCO2 and quality flags for a single year",
    x = "DOY", y = expression(FCO2~(g~C~m^{-2}~d^{-1})) 
  ) + my_theme
```
Where to go from here:
- Interpret the figure and include it in your group presentation.


# Optional: Ecosystem water budget
ET = Evapotranspiration = the sum of evaporation from soil + transpiration from plants.
P = Precipitation = all water input from rain/snow.
```{r}
# https://rdrr.io/cran/bigleaf/man/LE.to.ET.html
monthly_data <- monthly_data %>%
  mutate(y_label_pos = max(P, ET, na.rm = TRUE) * 1)

# Add ET/P ratio to monthly_data
monthly_data <- monthly_data %>%
  mutate(ET_P_ratio = ET / P)

# Add GPP/ET ratio to monthly_data for WUE
monthly_data <- monthly_data %>%
  mutate(WUE = -(GPP)/ET)

# monthly_data <- monthly_data %>%
#   mutate(y_label_pos = max(P, ET, na.rm = TRUE) * 1)

# plot: P and ET
ggplot(monthly_data, aes(x = month_year)) +
  geom_bar(aes(y = P, fill = "P"), stat = "identity", position = "dodge", alpha = 0.7) +
  geom_line(aes(y = ET, color = "ET"), size = 1) +
  geom_vline(data = monthly_data %>% filter(month == 12), 
             aes(xintercept = as.numeric(month_year) + 15), 
             linetype = "dashed", color = "black", size = 0.5) +
  geom_text(data = monthly_data %>% filter(month == 6), 
            aes(x = year_label_pos, 
                y = max(ET, P, na.rm = TRUE) * 1,  # Adjust label position based on max value
                label = year), 
            color = "black", size = 5, fontface = "bold") +
  labs(x = "Month", y = expression("mm of water"), title = "Monthly ET and P Trends") +
  scale_fill_manual(values = c("P" = "gray")) +  # P (bars)
  scale_color_manual(values = c("ET" = "#F8766D")) +  # ET (line)
  scale_x_date(date_labels = "%m", date_breaks = "6 month", expand = c(0, 0)) +  
  scale_y_continuous(
    breaks = seq(0, ceiling(max(monthly_data$ET, monthly_data$P, na.rm = TRUE) + 5), by = 50),
    limits = c(0, ceiling(max(monthly_data$ET, monthly_data$P, na.rm = TRUE) + 5))
  )  + my_theme

# # plot: monthly ET/T Ratio
# ggplot(monthly_data, aes(x = month_year, y = ET_P_ratio)) +
#   geom_bar(stat = "identity", fill = "#619CFF", alpha = 0.8) +
#   labs(x = "Month", y = "ET / P", title = "Monthly ET/P Ratio") +
#   geom_vline(data = monthly_data %>% filter(month == 12), 
#              aes(xintercept = as.numeric(month_year) + 15), 
#              linetype = "dashed", color = "black", size = 0.5) +
#   geom_hline(yintercept = 1, linetype = "dashed", color = "black", size = 0.8) +
#   geom_text(data = monthly_data %>% filter(month == 6), 
#             aes(x = year_label_pos, 
#                 y = max(ET/P, na.rm = TRUE) * 1.3,  # Adjust label position based on max value
#                 label = year), 
#             color = "black", size = 5, fontface = "bold") +
#   scale_x_date(date_labels = "%m", date_breaks = "6 month", expand = c(0, 0)) +
#   scale_y_continuous(
#     breaks = seq(0, ceiling(max(monthly_data$ET_P_ratio, na.rm = TRUE) + 0.5), by = 0.5),
#     limits = c(0, ceiling(max(monthly_data$ET_P_ratio, na.rm = TRUE) + 0.5))
#   ) + my_theme
#   
# plot: water use efficiency
ggplot(monthly_data, aes(x = month_year, y = WUE)) +
  geom_bar(stat = "identity", fill = "#619CFF", alpha = 0.8) +
  labs(x = "Month", y = "WUE", title = "Water-use efficiency (GPP/ET)") +
  geom_vline(data = monthly_data %>% filter(month == 12), 
             aes(xintercept = as.numeric(month_year) + 15), 
             linetype = "dashed", color = "black", size = 0.5) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "black", size = 0.8) +
  geom_text(data = monthly_data %>% filter(month == 6), 
            aes(x = year_label_pos, 
                y = 60,  # Adjust label position based on max value
                label = year), 
            color = "black", size = 5, fontface = "bold") +
  scale_x_date(date_labels = "%m", date_breaks = "6 month", expand = c(0, 0)) +
  scale_y_continuous(
    breaks = seq(0, ceiling(max(monthly_data$WUE, na.rm = TRUE) + 1), by = 10),
    limits = c(0, ceiling(max(monthly_data$WUE, na.rm = TRUE) + 0.5))
  ) + my_theme
# For winter, snowfall melt is 
# WUE: only for growing season: focus on peak growing season only ... why 
```
Where to go from here:
- Interpret the figure and include it in your group presentation.

# Optional: Light use efficiency
```{r}
data_filtered = df.HH[df.HH$month == 7, ] # specify the period you want like to focus on
data_filtered$PPFD = data_filtered$PPFD_IN
data_filtered$NEE = data_filtered$NEE_VUT_REF

# # # Function to calculate light response curve for NEE based on the Michaelis-Menten equation
# light_response_NEE <- function(PPFD, Amax, alpha, Rd) {
#   -((Amax * PPFD) / (alpha + PPFD) - Rd)
# }
# 
# 
# fit_NEE <- nls(NEE ~ light_response_NEE(PPFD, Amax, alpha, Rd), 
#                data = data_filtered, 
#                start = list(Amax = max(-data_filtered$NEE, na.rm = TRUE), 
#                             alpha = 200, Rd = 2))

# Extract model parameter estimates
params <- coef(fit_NEE)
Amax_est <- round(params["Amax"], 2)
alpha_est <- round(params["alpha"], 2)
Rd_est <- round(params["Rd"], 2)
# compute A2000
A2000 = Amax_est * 2000/(alpha_est + 2000)

# Generate predicted values from the fitted model
data_filtered$NEE_pred <- predict(fit_NEE, newdata = data_filtered)
ggplot(data_filtered, aes(x = PPFD, y = NEE))  +
  geom_point() +
  geom_line(aes(y = NEE_pred), color = "red", size = 1.2) +  
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +  
  geom_vline(xintercept = 1800, linetype = "dashed", color = "grey", size = 1) +
  labs(x = expression(PPFD ~ "(" * mu * "mol" ~ m^{-2} ~ s^{-1} * ")"), 
       y = expression(FCO[2] ~ "(" * mu * "mol" ~ m^{-2} ~ s^{-1} * ")")) +
  my_theme
  
fit_light_response_NEE <- function(data) {
  # Skip if not enough data
  if (nrow(data) < 10 || all(is.na(data$NEE))) return(NULL)

  # Michaelis-Menten light response model
  light_response_NEE <- function(PPFD, Amax, alpha, Rd) {
    -((Amax * PPFD) / (alpha + PPFD) - Rd)
  }

  # Try fitting the model, catch failures
  fit_result <- tryCatch({
    nls(NEE ~ light_response_NEE(PPFD, Amax, alpha, Rd),
        data = data,
        start = list(
          Amax = max(-data$NEE, na.rm = TRUE),  # guess based on data
          alpha = 200,
          Rd = 2
        ),
        control = nls.control(maxiter = 100, warnOnly = TRUE)
    )
  }, error = function(e) {
    message("Model failed for one year: ", e$message)
    return(NULL)
  })

  # If fitting was successful, extract A2000
  if (!is.null(fit_result)) {
    params <- coef(fit_result)
    Amax <- params["Amax"]
    alpha <- params["alpha"]
    A2000 <- Amax * 2000 / (alpha + 2000)
    return(data.frame(Amax = Amax, alpha = alpha, Rd = params["Rd"], A2000 = A2000))
  } else {
    return(NULL)
  }
}

# compute A2000 for each year
data_by_year <- split(data_filtered, data_filtered$year)
results_by_year <- lapply(data_by_year, fit_light_response_NEE)
names(results_by_year) <- names(data_by_year)
results_df <- bind_rows(results_by_year, .id = "Year")

ggplot(results_df, aes(x = as.numeric(Year), y = A2000)) +
  geom_line() +
  geom_point() +
  labs(title = "A2000 over Years", x = "Year", y = "A2000 (µmol m⁻² s⁻¹)") +
  ylim(0,50) + my_theme
```
Where to go from here:
- Interpret the figure and include it in your group presentation.
- Explore the seasonal pattern of A2000.
- Group discussion: What parameters used in land surface models relate to light use and photosynthesis? 


# Optional: Energy balance closure
```{r}

```
Where to go from here:
- Interpret the figure and include it in your group presentation.
- 



# Optional: canpy conductance
Using package bigleaf
canopy conductance for each month
pull GCC and integrate with flux data ... 
```{r}
library(bigleaf)
```
Where to go from here:
- Interpret the figure and include it in your group presentation.

