---
title: "flux-model-comparison"
output: html_document
date: "2025-08-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(data.table)
library(ggplot2)
library(ggpubr)
library(tidyverse)

my_theme <- theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(size = 12),
    # panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = "white")
  )

project_dir = ("G:/My Drive/FLUX-LSM workshop/US-Syv/")
site_name <- "US-Syv"
```
# read in flux data
```{r}
setwd(paste0(project_dir, "AMF_US-Syv_FLUXNET_SUBSET_2001-2023_4-6"))
# read in daily flux data
df.flux = fread("AMF_US-Syv_FLUXNET_SUBSET_DD_2001-2023_4-6.csv")
# covert -9999 to NA
df.flux <- df.flux %>%
  mutate(across(everything(), ~na_if(. , -9999)))

# add more time variables
df.flux <- df.flux %>%
  mutate(
    TIMESTAMP = ymd(TIMESTAMP),
    year = year(TIMESTAMP),
    month = month(TIMESTAMP),
    date = as.Date(TIMESTAMP),
    doy = yday(TIMESTAMP)
  )
```

# read in ELM model output
```{r}
library(ncdf4)
setwd(project_dir)
# Open the NetCDF file
nc <- nc_open("ELM_output_US_Syv.nc")

# Extract time (mcdate)
mcdate <- ncvar_get(nc, "mcdate")

# Extract selected variables
vars_to_extract <- c("NEE", # net ecosystem exchange of carbon
                     "GPP", # gross primary production
                     "ER", # ecosystem respiration
                     "TSOI_10CM", # soil temperature (10cm)
                     "FSH", # sensible heat
                     "EFLX_LH_TOT", # total latent heat flux
                     "RH2M") # 2m relative humidity
data_list <- lapply(vars_to_extract, function(var) ncvar_get(nc, var))

# Combine into data frame
df.model <- data.frame(mcdate = mcdate)
for (i in seq_along(vars_to_extract)) {
  df.model[[vars_to_extract[i]]] <- data_list[[i]]
}

# Close NetCDF file
nc_close(nc)

# modify time variables
df.model <- df.model %>%
  mutate(
    mcdate = ymd(mcdate),
    year = year(mcdate),
    month = month(mcdate),
    date = as.Date(mcdate),
    doy = yday(mcdate)
  )
df.model = df.model[df.model$year >=2001, ] # start year of flux measurement

df.combined = left_join(df.model, df.flux) # dataframe combining fluxnet variables with selected model output
names(df.combined)
```
# TSOI_10CM
## One to one plot
```{r}
# Flux measurement: TS_F_MDS_1, unit: °C
# Model output:  TSOI_10CM, unit: K
xlab = expression(Tsoil[flux]~(degree*C))
ylab = expression(Tsoil[model]~(degree*C))

df.combined$Tsoil_flux = df.combined$TS_F_MDS_1
df.combined$Tsoil_model = df.combined$TSOI_10CM - 273.15 # unit conversion

ggscatter(
  data = df.combined,
  x = "Tsoil_flux",  # flux measurement
  y = "Tsoil_model", # model variable after unit conversion
  color = "steelblue",
  add = "reg.line",
  add.params = list(color = "black", size = 1),  # regression line
  conf.int = TRUE,
  cor.coef = TRUE,
  cor.coeff.args = list(size = 6),
  label.x = min(df.combined$TS_F_MDS_1, na.rm = TRUE),
  label.y = max(df.combined$TS_F_MDS_1, na.rm = TRUE),
  title = "Tsoil Comparison",
  ggtheme = theme_pubr()
) +
stat_regline_equation(
  aes(label = ..eq.label..),
  label.x = -10, label.y = 18, size = 6
) + 
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed", size = 1) +
xlim(-15,25) + ylim(-15,25) +
  xlab(xlab) + ylab(xlab)
# setwd(project_dir); ggsave("figure_Tsoil.png") 
```
## Temporal comparison
```{r}
df_long <- df.combined %>%
  pivot_longer(cols = c(Tsoil_flux, Tsoil_model),
               names_to = "source",
               values_to = "Tsoil") %>%
  filter(year >= 2001 & year <= 2006)

ggplot(df_long, aes(x = doy, y = Tsoil, color = source)) +
  geom_smooth(se = FALSE, method = "loess", span = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ year, ncol = 2) +
  ylab(ylab) +
  scale_color_manual(values = c("Tsoil_flux" = "red", "Tsoil_model" = "blue"),
                     labels = c("Flux", "Model")) +
  my_theme
```
# NEE
## One to one plot
```{r}
# Flux measurement: NEE_VUT_REF, unit: gC m-2 d-1
# Model output: NEE, unit: gC/m2/s
df.combined$NEE_flux = df.combined$NEE_VUT_REF
# convert model variable to the same unit of fluxnet variable: gC m-2 d-1
df.combined$NEE_model <- df.combined$NEE * 86400 # number of seconds in a day
xlab = expression(NEE[flux]~'('*gC~m^{-2}~day^{-1}*')')
ylab = expression(NEE[model]~'('*gC~m^{-2}~day^{-1}*')')

ggscatter(
  data = df.combined,
  x = "NEE_flux",  # flux measurement
  y = "NEE_model",      # model variable after unit conversion
  color = "steelblue",
  add = "reg.line",
  add.params = list(color = "black", size = 1),  # regression line
  conf.int = TRUE,
  cor.coef = TRUE,
  cor.coeff.args = list(size = 6),
  label.x = min(df.combined$NEE_VUT_REF, na.rm = TRUE),
  label.y = max(df.combined$NEE_model, na.rm = TRUE),
  title = "NEE Comparison",
  ggtheme = theme_pubr()
) +
stat_regline_equation(
  aes(label = ..eq.label..),
  label.x = -10, label.y = 5, size = 6
) + 
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed", size = 1) +
  xlab(xlab) + ylab(ylab) +
  xlim(-10,10) + ylim(-10,10) 
```
## Temporal comparison
```{r}
#smooth line graph vs DOY
ylab = expression(NEE~'('*gC~m^{-2}~day^{-1}*')')
df_long <- df.combined %>%
  pivot_longer(cols = c(NEE_flux, NEE_model),
               names_to = "source",
               values_to = "flux") %>%
  filter(year >= 2001 & year <= 2006) # filter years of interest

ggplot(df_long, aes(x = doy, y = flux, color = source)) +
  geom_smooth(se = FALSE, method = "loess", span = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ year, ncol = 2) +
  ylab(ylab) +
  scale_color_manual(values = c("NEE_flux" = "red", "NEE_model" = "blue"),
                     labels = c("Flux", "Model")) +
  my_theme
```
## Cumulative NEE
```{r}
ylab = expression(Cumulative~NEE~(g~C~m^{-2}))
# Calculate cumulative NEE
df.cumulative <- df.combined %>%
  filter(!is.na(NEE_flux)) %>% # compare where both values exist
  arrange(mcdate) %>%
  mutate(
    cum_flux = cumsum(NEE_flux),
    cum_model = cumsum(NEE_model)
  )


# Plot cumulative NEE
ggplot(df.cumulative, aes(x = mcdate)) +
  geom_line(aes(y = cum_flux, color = "Flux")) +
  geom_line(aes(y = cum_model, color = "Model")) +
  labs(
    title = "Cumulative NEE",
    x = "Date",
    y = ylab,
    color = "Source"
  ) +
  scale_color_manual(values = c("Flux" = "blue", "Model" = "red")) +
  my_theme

# plot by cumulative NEE for each year
{
  df_cum_long <- df.combined %>%
  pivot_longer(cols = c(NEE_flux, NEE_model),
               names_to = "source", values_to = "flux") %>%
  group_by(year, source) %>%
  arrange(doy) %>%
  mutate(cumulative_flux = cumsum(replace_na(flux, 0))) %>%
  ungroup() %>%
  filter(year >= 2001 & year <= 2006)

df_labels <- df_cum_long %>%
  group_by(year, source) %>%
  filter(doy == max(doy)) %>%
  ungroup()

ggplot(df_cum_long, aes(x = doy, y = cumulative_flux, color = source, group = source)) +
  geom_line(size = 1) +
  facet_wrap(~ year, ncol = 2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey") +
  geom_text(data = df_labels,
            aes(label = round(cumulative_flux, 1)),
            hjust = -0.1, vjust = 0.5,
            size = 3, show.legend = FALSE) +
  scale_color_manual(values = c("NEE_flux" = "blue", "NEE_model" = "red"),
                     labels = c("Flux", "Model")) +
  labs(
    title = "cumulative NEE",
    x = "Day of Year (DOY)",
    y = "Cumulative NEE by year (gC m⁻²)",
    color = "Source"
  ) + my_theme

}
```
## Monthly sums
```{r}
df_monthly_sum <- df.combined %>%
  filter(year >= 2001 & year <= 2006) %>%
  group_by(year, month) %>%
  summarise(
    NEE_flux = sum(NEE_flux, na.rm = TRUE),
    NEE_model = sum(NEE_model, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(NEE_flux, NEE_model),
               names_to = "source",
               values_to = "monthly_sum")

ggplot(df_monthly_sum, aes(x = month, y = monthly_sum, fill = source)) +
  geom_col(position = "dodge") +
  facet_wrap(~ year, ncol = 2) +
  scale_fill_manual(
    values = c("NEE_flux" = "blue", "NEE_model" = "red"),
    labels = c("Flux", "Model")
  ) +
  scale_x_continuous(breaks = 1:12) +
  labs(
    title = "Monthly NEE",
    x = "Month",
    y = "NEE Monthly (gC m⁻²)",
    fill = "Data Source"
  ) + my_theme
```



# GPP
## one to one plot
```{r}
# FLUX measurement: GPP_VUT_REF, unit: gC m-2 d-1
# Model output: GPP, unit: gC/m2/s

df.combined$GPP_flux = df.combined$GPP_NT_VUT_REF
# convert model variable to the same unit of fluxnet variable: gC m-2 d-1
df.combined$GPP_model <- df.combined$GPP * 86400 # number of seconds in a day
xlab = expression(GPP[flux]~'('*gC~m^{-2}~day^{-1}*')')
ylab = expression(GPP[model]~'('*gC~m^{-2}~day^{-1}*')')
ggscatter(
  data = df.combined,
  x = "GPP_flux", # flux measurement
  y = "GPP_model", # model variable after unit conversion
  color = "darkgreen",
  add = "reg.line",
  add.params = list(color = "black", size = 1),  # regression line 
  conf.int = TRUE,
  cor.coef = TRUE,
  cor.coeff.args = list(size = 6),  
  label.x = min(df.combined$GPP_NT_VUT_REF, na.rm = TRUE),
  label.y = max(df.combined$GPP, na.rm = TRUE),
  xlab = "GPP_NT_VUT_REF (Flux)",
  ylab = "GPP (Model)",
  title = "GPP Comparison",
  ggtheme = theme_pubr()
) +
stat_regline_equation(
  aes(label = ..eq.label..),
  label.x = 0, label.y = 15, size = 6
) + 
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed", size = 1) +
  xlab(xlab) + ylab(ylab) + 
  xlim(0,15) + ylim(0,15)

```
# Evaluating level of agreement
NOTE: this is a simplified approach that evaluates model-data agreement assuming model and obs data are from a single point, over the entire period that data is available for both, instead of looking at specific intervals where data is available and integrating over those for the full record, we just dropped time periods where there isn't observational data to compare to the model. This means that model-data agreement isn't scored for interannual variability, and there aren't spatial components to the evaluation. 
## subset of data
```{r}
variables_compare <- c("year", "month", "date", "doy", "TIMESTAMP", "NEE_flux", "NEE_model", "GPP_flux", "GPP_model")
df.combined_sub <- df.combined[,variables_compare]
#replace negative tower GPP values w/NA for now
df.combined_sub$GPP_flux[df.combined_sub$GPP_flux < 0] <- NA


df.combined_sub <- df.combined_sub %>%
  pivot_longer(cols = c(NEE_flux, NEE_model, GPP_flux, GPP_model),
             names_to = "variable",
             values_to = "value")

df.combined_sub <- df.combined_sub %>%
  mutate(data_type = case_when(variable == "NEE_flux" ~ "Tower",
                               variable == "NEE_model" ~ "Model",
                               variable == "GPP_flux" ~ "Tower",
                               variable == "GPP_model" ~ "Model",))

df.combined_sub <- df.combined_sub %>%
  mutate(flux_type = case_when(variable == "NEE_flux" ~ "NEE",
                               variable == "NEE_model" ~ "NEE",
                               variable == "GPP_flux" ~ "GPP",
                               variable == "GPP_model" ~ "GPP",))

df.combined_sub$site_name <- site_name
```
## filter data
```{r}
#filtering out dates where there isn't tower data available for comparison
# Identify dates where Tower data is missing for each flux_type
missing_dates <- df.combined_sub %>%
  filter(data_type == "Tower") %>%
  filter(is.na(value)) %>%
  select(date, flux_type)

# Remove all rows for those dates and flux_type combos
df.combined_sub <- df.combined_sub %>%
  anti_join(missing_dates, by = c("date", "flux_type"))
```
## compute metrics
```{r}
# Variables to evaluate
variable_list <- c("GPP", "NEE")
all_metrics <- list()  # to store results

for (variable_run in variable_list) {
  
  # Subset model and observation data
  model <- subset(df.combined_sub, data_type == "Model" & flux_type == variable_run)
  obs <- subset(df.combined_sub, data_type == "Tower" & flux_type == variable_run)
  
  norm_sd <- sd(model$value) / sd(obs$value)
  
  # Correlation score -- tells us if the model gets the general temporal pattern right
  S_corr <- (1 + cor(model$value, obs$value, method = "pearson")) / 2
  
  # Bias and bias score -- tells us if the model consistently over or underestimates
  bias <- model$value - obs$value #pointwise 
  
  #nondimensionalize bias as a relative error, account for cases where you might be dividing by 0 b/c obs values could be 0
  e_bias <- ifelse(
    abs(obs$value) > 0,
    abs(model$value - obs$value) / abs(obs$value),
    NA)
  
  S_bias <- exp(-1 * e_bias)
  
  S_bias[!is.finite(S_bias)] <- NA
  
  domain_S_bias <- mean(S_bias, na.rm = TRUE)
  
  # RMSE and RMSE score -- tells us how well the amplitude and variability match (this approach controls for double-counting bias by using centralized rmse)
  #This implementation assumes no spatial structure (i.e., it’s a single grid cell or averaged across an area of interest). 
  #If you're computing this per grid cell in a spatial array, you'd apply it per column or group
  #This approach removes bias before computing RMSE, so the RMSE score focuses on variance differences rather than 
  #systematic offset, and so bias doesn't get essentially 'double counted' in the overall score
  #first need to centralize (remove mean) the timeseries
  model_anom <- model$value - mean(model$value, na.rm=T)
  obs_anom <- obs$value - mean(obs$value, na.rm=T)
  
  #then compute centralized rmse
  n <- length(obs_anom)
  crmse <- sqrt(mean((model_anom - obs_anom)^2, na.rm=T))
  
  #and compute centralized RMS of reference obs (crms)
  crms <- sqrt(mean(obs_anom^2, na.rm = TRUE))
  
  # calculate relative error
  e_rmse <- crmse / crms
  
  S_rmse <- exp(-1 * e_rmse)
  
  # Overall score
  S <- (domain_S_bias + 2 * S_rmse + S_corr) / 4
  
  # Summary metrics
  metric <- c("tower mean", "tower SD","model mean", "model SD", "SD norm.", "bias", "RMSE", 
              "correlation score", "bias score", "RMSE score", "overall score")
  
  model_scores <- c(mean(obs$value), sd(obs$value), mean(model$value), sd(model$value), norm_sd,
                     mean(bias), rmse, S_corr, domain_S_bias, S_rmse, S)
  
  
  score_df <- data.frame(site = site_name, variable = variable_run, metric = metric, value = model_scores)
  
  all_metrics[[variable_run]] <- score_df
}

# Combine results from loop
metric_df <- do.call(rbind, all_metrics)
row.names(metric_df) <- NULL

print(metric_df) #check out the results!

```
## plot the results
```{r}
library(RColorBrewer)
# plot subset of the metrics
scores <- c("correlation score", "bias score", "RMSE score", "overall score") 
metric_df_sub <- subset(metric_df, metric_df$metric %in% scores)
var_list <- c("GPP", "NEE")
metric_df_sub <- subset(metric_df_sub, metric_df_sub$variable %in% var_list)
metric_df_sub$metric <- factor(metric_df_sub$metric, 
                               levels = c("bias score", "correlation score", 
                                          "RMSE score", "overall score"))


ggplot(metric_df_sub, aes(x = metric, y = site, fill = value)) +
  geom_tile(color = "white") +
  geom_text(aes(label = sprintf("%.2f", value)), size = 5) +  # Display metric values
  scale_fill_gradientn(colors = brewer.pal(n = 10, name = "Spectral"), limit = c(0, 1), name = "Metric Score") +
  facet_wrap(~ variable, scales = "free", ncol = 3) +
  my_theme + 
  labs(title = NULL,
       x = "Evaluation Metric",
       y = NULL) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

#plot_save <- "model_scoring_GPP_NEE.png"
#ggsave(filename = plot_save, dpi = 500)
```






